---
title: "Assignment 2 Q1"
author: "Zhen Xien Lim"
date: "5 September 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Question 1
We first store the dates in vector form.
```{r}
A=as.Date(c("17march1964","20june1992",""),"%d%B%Y")
B=as.Date(c("06may1964","06august1992","12june1993"),"%d%B%Y")
C=as.Date(c("12august1964","18december1992","18june1995"),"%d%B%Y")
D=as.Date(c("27october1964","04january1993",""),"%d%B%Y")
E=as.Date(c("04january1865","28april1993","29august1996"),"%d%B%Y")
F=as.Date(c("18april1965","16june1993","12december1995"),"%d%B%Y")
G=as.Date(c("20may1965","29october1993","21april1996"),"%d%B%Y")
H=as.Date(c("04july1965","16february1994",""),"%d%B%Y")
I=as.Date(c("16september1965","22august1994",""),"%d%B%Y")
J=as.Date(c("11december1965","06march1995","17february1997",""),"%d%B%Y")
```
To identify the central exposed to risk time for each life, we take the latest of date reaching   30, start of investigation and date of entr for Date A, earliest of date reaching age 31, end of  investigation and date of exit for Date B. Then the difference of Date A and Date B will be the   central exposed to risk. I will assume withdrawal of policy date to be at the end of the policy anniversary year. 
```{r}
life=c()
life[1]=(((A[1]+(31*365.25))-(A[1]+(30*365.25)))/365.25) ## one year of exposed to risk
life[2]=0  #the date of exit is earlier than the date of reaching age 30, therefore life B is not exposed to risk.
life[3]=(((C[1]+(31*365.25))-(C[1]+(30*365.25)))/365.25)
life[4]=(((D[1]+(31*365.25))-(D[1]+(30*365.25)))/365.25)
life[5]=0
life[6]=(F[3]-(F[1]+(30*365.25)))/365.25
life[7]=(G[3]-(G[2]+(2*365.25)))/365.25
life[8]=(((H[1]+(31*365.25))-(H[1]+(30*365.25)))/365.25)
life[9]=1    
life[10]=(((J[3]-(J[2]+365.25))/365.25))
q30=2/sum(life)
q30
```

```{r}
lifem=c()
lifem[1]=(((A[1]+(31*365.25))-(A[1]+(30*365.25)))/365.25) ## one year of exposed to risk
lifem[2]=0  #the date of exit is earlier than the date of reaching age 30, therefore life B is not exposed to risk.
lifem[3]=(C[3]-(C[1]+(30*365.25)))/365.25
lifem[4]=(((D[1]+(31*365.25))-(D[1]+(30*365.25)))/365.25)
lifem[5]=0
lifem[6]=(F[3]-(F[1]+(30*365.25)))/365.25
lifem[7]=(G[3]-(G[2]+(2*365.25)))/365.25
lifem[8]=(((H[1]+(31*365.25))-(H[1]+(30*365.25)))/365.25)
lifem[9]=0     
lifem[10]=(((J[3]-(J[2]+365.25))/365.25))
```

We then calculate the MLE of q30 under assumption of constant force of mortality by deaths divided by the central exposed to risk. 
```{r}
MLEq30=2/sum(lifem)
MLEq30
```
