---
title: "Assignment 2 Question 2"
author: "Zhen Xien Lim"
date: "5 September 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Question 2
```{r}
agegroup=c(1:11)
EtR=c(15518,19428,21594,21890,19174,15775,11414,6993,3276,1096,2001)
Actdeath=c(65,144,219,378,465,557,685,644,471,217,67)
totaldeath=sum(Actdeath)
Expdeath=c(73.9,134.6,223.9,346.3,468.1,600.2,675.5,637.4,458.7,240.6,61.4)
qx=Expdeath/EtR
Vardeath=Expdeath*(1-qx)
```
a) Chi-square test
```{r}
Zx=(Actdeath-Expdeath)/sqrt(Vardeath)
chi_2=Zx^2
sumchi_2=sum(chi_2)
qchisq(.95,11)
```
chi-square value witht 11 degree of freedom is higher than the sum of all chi-square with 
1 degree freedom each. Therefore we cannot reject the null hypothesis. We can also calculate 
the p-value.
However, the problem of aggregate chi-square distribution is that it does not tell the
individual behaviour of the data.
```{r}
pchisq(sumchi_2,11)
```

b) Standard Deviation test
Fot the Z-scores, we ecpect to have the ranges of (-3,-2),(-2,-1),(-1,0),(0,1),(1,2),(2,3)
and we generate the probabilities of being in these ranges.
```{r}
r=c()
r[1]=pnorm(-3)
r[2]=pnorm(-2)-pnorm(-3)
r[3]=pnorm(-1)-pnorm(-2)
r[4]=pnorm(0)-pnorm(-1)
r[5]=pnorm(1)-pnorm(0)
r[6]=pnorm(2)-pnorm(1)
r[7]=pnorm(3)-pnorm(2)
r[8]=1-pnorm(3)
r
```
Now we calculate the expected numbers falling in these ranges.
```{r}
Er=11*r
Er
```

```{r}
Ar=c()
Ar[1]=sum(Zx<(-3))
Ar[2]=sum((Zx>(-3))*(Zx<(-2)))
Ar[3]=sum((Zx>(-2))*(Zx<(-1)))
Ar[4]=sum((Zx>(-1))*(Zx<(0)))
Ar[5]=sum((Zx>(0))*(Zx<(1)))
Ar[6]=sum((Zx>(1))*(Zx<(2)))
Ar[7]=sum((Zx>(2))*(Zx<(3)))
Ar[8]=sum(Zx>(3))
Ar

t1=c()
t1[1]=(sum(Ar[1:3])-sum(Er[1:3]))^2/sum(Er[1:3])
t1[2]=(Ar[4]-Er[4])^2/Er[4]
t1[3]=(Ar[5]-Er[5])^2/Er[5]
t1[4]=(sum(Ar[6:8])-sum(Er[6:8]))^2/sum(Er[6:8])
t1
```
The p-value of the chi-square statistics with dof 3 is not significant at the 0.5% level.
Therefore we can say taht the S.D distribution is unit normal.
```{r}
1-pchisq(sum(t1),3)
```
The probability that a unit normal lies between 2/3 and 02/3 is almost 0.5.
We then count the absolute Z-scores(SD) that are greater than 2/3
```{r}
pnorm(2/3)-pnorm(-2/3)
```

```{r}
AbDev=sum(abs(Zx)>2/3)
AbDev
```

Using a binomial (11,0.5) distribution, we can work out what the probability of such a count is.
it is not significant, which gives not enough evidence to reject the null hypothesis, which 
is the data used to calculate the expected death is suitable.
```{r}
1-pbinom(AbDev,11,0.5)
```
c) Cumulative Deviation test
it is smaller than the 1.96 absolute value = therefore we can say that the adherence to data 
is good.
```{r}
CV=(sum(Actdeath)-sum(Expdeath))/sqrt(sum(Vardeath))
CV
``` 
d) Serial Correlation test
The test result is smaller than the 1.64 at CI95%, so we cannot reject the null hypothesis, the adherence of data is good. 
```{r}
zi=Zx[1:10]
ziplus1=Zx[2:11]
z1=sum(zi)/10
z2=sum(ziplus1)/10
test=sum((zi-z1)*(ziplus1-z2))/sqrt((sum(zi-z1))^2*(sum(ziplus1-z2))^2)
test
```
e) Sign test
```{r}
S=length(Zx)
i=1
V=0
while (i<=S) {
  if(Zx[i]>0){
    V=V+1
  }
  i=i+1
}
print(V)
```

